Gromov--Witten theory of hypersurfaces in genus one

### 2019-02-08 GenusOneRelativeMaps.tex
Finished description of rubber recursion

\subsection{Recursion procedure for reduced rubber integrals} In the recursion above, certain terms appear which are integrals over moduli spaces of reduced log stable maps to rubber (as defined in \S \ref{}):
\begin{equation*}\VZ_{1,\alpha}(P|H_0+H_\infty)_{\Gm}.\end{equation*}
In this section we provide a recursive algorithm for computing such integrals. For convenience we denote the \textbf{horizontal degree} of the rubber map by:
\begin{equation*} d := \sum_{i=1}^n \alpha_i .\end{equation*}
Recall that the $\alpha_i$ may be positive or negative (recording tangencies with respect to $H_\infty$ and $H_0$, respectively) and that $d \geq 0$ always. First suppose that $d=0$. Then the usual rubber moduli space decomposes as
\begin{equation*} \MLog_{1,\alpha}(P|H_0 + H_\infty)_{\Gm} = \MLog_{1,\alpha}(\PP^1|0+\infty)_{\Gm} \times H \end{equation*}
and the first term on the right-hand side is birational to the double ramification locus in $\ol\Mcal_{1,n}$. This locus has a main component (the closure of the locus of smooth curves) and one other ``exceptional'' component, consisting of split curves where all the markings belong to the rational tail. The corresponding reduced rubber space is birational to the main component, and thus we can identify its class by subtracting off the class of the exceptional component. This gives
\begin{equation*} \pi_\st [\VZ_{1,\alpha}(P|H_0+H_\infty)_{\Gm}] = \left( [\ol\Mcal_{1,n}^\alpha] - [D] \right) \times [H] \in \Achow_\st ( \ol\Mcal_{1,n} \times H ) = \Achow_\st ( \ol\Mcal_{1,n}(H,0) ) \end{equation*}
where $\pi \colon \VZ_{1,\alpha}(P|H_0+H_\infty)_{\Gm} \to \ol\Mcal_{1,n}(H,0)$ is the map which forgets the log structure and collapses the map, $D$ is the boundary locus of split curves described above, and $[\ol\Mcal_{1,n}^\alpha]$ is the classical double ramification cycle, which has an explicit formula in terms of tautological classes \cite{Hain,JPPZ}. This allows us to compute the integral over the rubber space as a tautological integral on the moduli space of curves.

Now suppose that $d>0$. If there is a marking $x_k$ which carries no insertion and has $\alpha_k \neq 0$, then consider the morphism which forgets $x_k$ and collapses the map:
\begin{equation*} \VZ_{1,\alpha}(P|H_0+H_\infty)_{\Gm} \xrightarrow{\pi} \VZ_{1,n-1}(H,d). \end{equation*}
These moduli spaces are log smooth and of the same dimension. Over the nice locus in $\VZ_{1,n-1}(H,d)$, $\pi$ has degree $\alpha_k^2$, arising from a choice of $\alpha_k$-torsion point on the source elliptic curve. Thus we have:
\begin{equation*} \pi_\st [\VZ_{1,\alpha}(P|H_0+H_\infty)_{\Gm}] = \alpha_k^2\cdot [\VZ_{1,n-1}(H,d)].\end{equation*}
Since all the insertions are pulled back from the right-hand side (up to modifying psi classes by some boundary strata which we can then also compute recursively) we can perform the integral on $\VZ_{1,n-1}(H,d)$ inductively.\footnote{(Navid) Say something here about the degree being smaller}

It remains to deal with the case where all of the relative markings carry insertions. The idea is to reduce to the case above by introducing an additional marking. Note first that every reduced rubber integral which appears in our recursion must either have a marking of negative tangency (i.e. tangent to $H_0$), or a marking with no insertions: indeed, the only boundary locus with no negative-tangency marking occurs when the entire curve is mapped to higher level, in which case (since we are assuming $d>0$) the curve is generically irreducible, and on this curve we have at least one fictitious marking (of positive tangency $1$) with no insertion.

@@

Now choose a point $x_1$ which carries only evaluation classes, and let $x_0$ denote the new non-relative marking we have introduced. If $\gamma$ was the class we were originally trying to integrate, consider the class $\tilde\gamma$ on \eqref{bigger rubber space} obtained from $\gamma$ by replacing $\ev_1$ by $\ev_0$. We obtain a recursion formula:
\begin{equation*} (\Psi_0+\Psi_\infty)\cdot \tilde\gamma \cap [\VZ_{1,\alpha\cup(0)}(P|H_0+H_\infty)_{\Gm}] = \sum_{\Gamma} m_\Gamma \cdot \tilde\gamma \cap [\ol\Mcal_{\Gamma}]. \end{equation*}
The right-hand side is a sum over irreducible components $\ol\Mcal_\Gamma$ of the locus where the target splits, with multiplicites $m_\Gamma$ calculated from the tropical picture. Each of these components can be written as a fibre product of (reduced) rubber spaces. One of these components is the locus where $x_0$ and $x_1$ belong to a trivial rational bubble:

@@

The corresponding moduli space is $\VZ_{1,\alpha}(P|H_0+H_\infty)_{\Gm}$ and we have $\ev_0=\ev_1$ here, so the contribution of this locus is precisely the original rubber integral we were trying to calculate. The remaining terms on the right-hand side have, by design, no insertion at $x_1$ and hence can be computed via the trick described previously. Finally, the left-hand side also has no insertion at $x_1$ and so can be computed by the same method. Here we are assuming that $x_1$ is not the only negative-tangency marking, so that $\Psi_0$ can be interpreted as an insertion on a marking different from $x_1$; if this is not the case, then a similar argument works by splitting $x_1$ into two markings of tangency $(\alpha_1+1,-1)$.

\textbf{What if $\alpha_1=-1$?}

@@

\textbf{Addendum} You have to be a bit careful about the target psi class on the larger space. We need to integrate over this, and the standard approach is to identify it with $\alpha_i \psi_i + \ev_i^\st H$ for $x_i$ any marking tangent to the corresponding divisor. Remember we are trying to move all the insertions off $x_k$. If there is some other marking $x_i$ tangent to the same divisor as $x_k$ then we're fine - just interpret the target psi class as an insertion on $x_i$. If not, we're forced to get an insertion at $x_k$ on the larger space, and we're stuck. In this case, we need to modify the recursion scheme slightly: instead of introducing an additional non-relative marking, split $x_k$ into two markings of tangency $1$ and $\alpha_k - 1$ (really, any splitting of $\alpha_k$ will probably do). Then we can safely move all the insertions (including the target psi class) onto one of these, and leave the other free to forget. Note this works even if $\alpha_k=1$.\medskip

>>>>

\subsection{Recursion procedure for reduced rubber integrals} In the recursion above, certain terms appear which are integrals over moduli spaces of reduced log stable maps to rubber (as defined in \S \ref{}):
\begin{equation*}\VZ_{1,\alpha}(P|H_0+H_\infty)_{\Gm}.\end{equation*}
In this section we provide a recursive algorithm for computing such integrals. We have three main techniques at our disposal: direct comparison with the double ramification cycle, forgetting of a marking with no insertion, and rubber recursion. Each technique has its own domain of applicability, and these domains have non-trivial intersection. For convenience we denote the \textbf{horizontal degree} of the rubber map by:
\begin{equation*} d := \sum_{i=1}^n \alpha_i .\end{equation*}
Recall that the $\alpha_i$ may be positive or negative (recording tangencies with respect to $H_\infty$ and $H_0$, respectively) and that $d \geq 0$ always. First suppose that $d=0$. Then the usual rubber moduli space decomposes as
\begin{equation*} \MLog_{1,\alpha}(P|H_0 + H_\infty)_{\Gm} = \MLog_{1,\alpha}(\PP^1|0+\infty)_{\Gm} \times H \end{equation*}
and the first term on the right-hand side is birational to the double ramification locus in $\ol\Mcal_{1,n}$. This locus has a main component (the closure of the locus of smooth curves) and one other ``exceptional'' component, consisting of split curves where all the markings belong to the rational tail. The corresponding reduced rubber space is birational to the main component, and thus we can identify its class by subtracting off the class of the exceptional component. This gives
\begin{equation*} \pi_\st [\VZ_{1,\alpha}(P|H_0+H_\infty)_{\Gm}] = \left( [\ol\Mcal_{1,n}^\alpha] - [D] \right) \times [H] \in \Achow_\st ( \ol\Mcal_{1,n} \times H ) = \Achow_\st ( \ol\Mcal_{1,n}(H,0) ) \end{equation*}
where $\pi \colon \VZ_{1,\alpha}(P|H_0+H_\infty)_{\Gm} \to \ol\Mcal_{1,n}(H,0)$ is the map which forgets the log structure and collapses the map, $D$ is the boundary locus of split curves described above, and $[\ol\Mcal_{1,n}^\alpha]$ is the classical double ramification cycle, which has an explicit formula in terms of tautological classes \cite{Hain,JPPZ}. These observations allow us to compute the integral over the rubber space as a tautological integral on the moduli space of curves.

Now suppose that $d>0$. If there is a marking $x_k$ which carries no insertion and has $\alpha_k \neq 0$, then consider the morphism which forgets $x_k$ and collapses the map:
\begin{equation*} \VZ_{1,\alpha}(P|H_0+H_\infty)_{\Gm} \xrightarrow{\pi} \VZ_{1,n-1}(H,d). \end{equation*}
The moduli spaces above are log smooth and of the same dimension. Over the nice locus in $\VZ_{1,n-1}(H,d)$, $\pi$ has degree $\alpha_k^2$, arising from a choice of $\alpha_k$-torsion point on the source elliptic curve. Thus we have:
\begin{equation*} \pi_\st [\VZ_{1,\alpha}(P|H_0+H_\infty)_{\Gm}] = \alpha_k^2\cdot [\VZ_{1,n-1}(H,d)].\end{equation*}
Since all the insertions are pulled back from the right-hand side (up to modifying psi classes by some boundary strata which we can then also compute recursively) we can perform the integral on $\VZ_{1,n-1}(H,d)$ inductively. Notice that, when we apply our recursion \ref{} to compute a reduced invariant of $H$, the only non-trivial reduced rubber integrals which appear will have smaller degree than the one we started with, so the calculation is not circular.

It remains to deal with the case where all of the relative markings carry insertions. The idea is to reduce to the case above by introducing an additional marking. Note first that every reduced rubber integral which appears in our recursion must either have a marking of negative tangency (i.e. tangent to $H_0$), or a marking with no insertions: indeed, the only boundary loci with no negative-tangency markings have the entire curve mapped to higher level, in which case (since we are assuming $d>0$, so we are not in the $C_0$ case) the curve is generically irreducible, and on this curve we have at least one fictitious marking (of positive tangency $1$) with no insertion.

@@

Now choose a negative-tangency point $x_1$ which carries only evaluation classes, and let $x_0$ denote the new non-relative marking we have introduced. If $\gamma$ was the class we were originally trying to integrate, consider the class $\tilde\gamma$ on \eqref{bigger rubber space} obtained from $\gamma$ by replacing $\ev_1^\st$ by $\ev_0^\st$. We obtain a recursion formula:
\begin{equation*} (\Psi_0+\Psi_\infty)\cdot \tilde\gamma \cap [\VZ_{1,\alpha\cup(0)}(P|H_0+H_\infty)_{\Gm}] = \sum_{\Gamma} m_\Gamma \cdot \tilde\gamma \cap [\ol\Mcal_{\Gamma}]. \end{equation*}
The right-hand side is a sum over irreducible components $\ol\Mcal_\Gamma$ of the locus where the target splits, with multiplicites $m_\Gamma$ calculated from the tropical data. Each of these components can be written as a fibre product of (reduced) rubber spaces. One of these components is the locus where $x_0$ and $x_1$ belong to a trivial rational bubble:

@@

The corresponding moduli space is $\VZ_{1,\alpha}(P|H_0+H_\infty)_{\Gm}$ and we have $\ev_0=\ev_1$ here, so the contribution of this locus is precisely the original rubber integral we were trying to calculate. The remaining terms on the right-hand side can be written either as a product of rubber integrals of smaller degree (which we can compute recursively), or rubber integrals of the same degree but with no insertion at $x_1$ (which we can compute by forgetting $x_1$). Finally, the left-hand side also has no insertion at $x_1$ and so can be computed by the same method. Here we are assuming that $x_1$ is not the only negative-tangency marking, so that $\Psi_0$ can be interpreted as an insertion on a marking different from $x_1$.

If $x_1$ is in fact the only negative-tangency marking, then we can modify the above argument by considering the following rubber space of dimension two larger:
\begin{equation}\label{rubber space two extra markings} \VZ_{1,\alpha\cup(0,0)}(P|H_0+H_\infty)_{\Gm} \end{equation}
Let us denote the newly-introduced non-relative markings by $x_0$ and $x_{n+1}$. We now let $\tilde\gamma$ denote the class obtained from $\gamma$ by replacing $\ev_1^\st$ by $\ev_0^\st$ and then multiplying by $\psi_{n+1}$. Following \cite{EKatz}\footnote{(Navid) Tropica interpretation of this? Vanishing multiplicities?}, there is a line bundle with a section cutting out the locus where $x_{n+1}$ does not belong to the leftmost component of the target. The first Chern class of this line bundle is $\Psi_\infty+\ev_{n+1}^*H$, which we may multiply with $\tilde\gamma$ and integrate over \eqref{rubber space two extra markings} to obtain a rubber integral with no insertion at $x_1$ (which we may then compute by forgetting $x_1$). On the right-hand side we obtain the integral of $\tilde\gamma$ over the loci where $x_{n+1}$ does not belong to the leftmost component of the target (this is just a union of some of the irreducible components of the split locus described above). One such locus is the following
[FIGURE]
and integrating $\tilde\gamma$ over this locus gives
\begin{equation*} \gamma\cdot\psi_{n+1}\cap [\VZ_{1,\alpha\cup(0)}(P|H_0+H_\infty)_{\Gm}] \end{equation*}
which we may identify with a (nonzero) multiple of our original rubber integral via the rubber dilaton equation (see\cite{MaulikPandharipande}; the same proof carries over to the reduced setting\footnote{(Navid) Think about this a bit}). The remaining loci may be computed recursively exactly as before, and thus we obtain a complete recursion procedure.

\begin{remark} This idea --- of finding a section of a line bundle cutting out the locus of split curves, and using this to derive recursion formulae for rubber integrals --- has been pursued in the non-reduced setting in \cite{EKatz}.\end{remark}

@@

\textbf{Addendum} You have to be a bit careful about the target psi class on the larger space. We need to integrate over this, and the standard approach is to identify it with $\alpha_i \psi_i + \ev_i^\st H$ for $x_i$ any marking tangent to the corresponding divisor. Remember we are trying to move all the insertions off $x_k$. If there is some other marking $x_i$ tangent to the same divisor as $x_k$ then we're fine - just interpret the target psi class as an insertion on $x_i$. If not, we're forced to get an insertion at $x_k$ on the larger space, and we're stuck. In this case, we need to modify the recursion scheme slightly: instead of introducing an additional non-relative marking, split $x_k$ into two markings of tangency $1$ and $\alpha_k - 1$ (really, any splitting of $\alpha_k$ will probably do). Then we can safely move all the insertions (including the target psi class) onto one of these, and leave the other free to forget. \medskip
